---
title: 'CRISPR sgRNA Library exploratory analysis: Kusa v3'
author: "Alex Kalinka"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    depth: 3
    highlight: tango
    number_sections: true
    theme: spacelab
    toc: true
    toc_float: true
---


```{r,echo=F}
options(warn=-1)

suppressMessages(library(knitr))
suppressMessages(library(tidyr))
suppressMessages(library(magrittr))
suppressMessages(library(readr))
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))

base_path.data <- file.path("~","data","az_cruk")
base_path.ht29 <- file.path(base_path.data,"ht29")

```

# Per-gene sgRNA counts

```{r}
# Count sgRNAs per gene.
count_sgrnas <- function(file, rename){
  if(rename){
    counts <- read.table(file, stringsAsFactors = F) %>%
      rename(sgRNA = V1, gene = V2)
  }else{
    counts <- read.table(file, header = T, stringsAsFactors = F)
  }
  counts %<>%
    filter(gene != "Control" & gene != ".") %>%
    group_by(gene) %>%
    summarise(sgRNA_per_gene_count = n()) %>%
    ungroup() %>%
    arrange(gene)
  return(counts)
}

# Plot sgRNA per-gene count histograms.
plot_sgrna_hist <- function(data, source){
  pl <- data %>%
    ggplot(aes(sgRNA_per_gene_count)) +
    geom_histogram() +
    xlim(0,10) +
    ggtitle(paste("sgRNA per-gene counts:",source))
  suppressMessages(print(pl))
}

# Summarise count distribution.
summarise.sgrna_counts <- function(data, name){
  data %<>%
    do(data.frame(data = "sgRNA_per_gene_counts", source = name,
              statistic = c("Min","Median","Max","Percent genes <6 sgRNAs",
                          "Percent genes >6 sgRNAs",
                          "Number genes 1 sgRNA"),
              value = c(min(.$sgRNA_per_gene_count, na.rm=T),
                            median(.$sgRNA_per_gene_count, na.rm=T),
                            max(.$sgRNA_per_gene_count, na.rm=T),
                            100*length(which(.$sgRNA_per_gene_count<6))/nrow(.),
                            100*length(which(.$sgRNA_per_gene_count>6))/nrow(.),
                            length(which(.$sgRNA_per_gene_count==1))))
    ) %>%
    mutate(value = round(value))
  return(data)
}

```

Both the plasmid count files from individual experiments and the Kusa v3 library exhibit the same distribution of per-gene sgRNA counts indicating that ~17% of the genes have fewer than 6 sgRNAs, ~4% of the genes have more than 6 sgRNAs, and 431 genes have exactly 1 sgRNA associated with them.

## HT29 {.tabset}

### Plasmid: Sanger

```{r}
counts_plasmid.sanger <- count_sgrnas(file.path(base_path.ht29,
                                                "crisprn","plasmid_test","counts","Plasmid_Sanger.s_1.counts"),
                                      rename=T)

plot_sgrna_hist(counts_plasmid.sanger, "HT29; Plasmid Sanger")

```

### Plasmid: replicate 1

```{r}
counts_plasmid.1 <- count_sgrnas(file.path(base_path.ht29,
                                                "crisprn","plasmid_test","counts","Plasmid_AZ_1.s_1.counts"), 
                                 rename=T)

plot_sgrna_hist(counts_plasmid.1, "HT29; Plasmid Replicate 1")

```

### Plasmid: replicate 2

```{r}
counts_plasmid.2 <- count_sgrnas(file.path(base_path.ht29,
                                                "crisprn","plasmid_test","counts","Plasmid_AZ_2.s_1.counts"), 
                                 rename=T)

plot_sgrna_hist(counts_plasmid.2, "HT29; Plasmid Replicate 2")

```

### Control: post-selection

```{r}
counts_ps <- count_sgrnas(file.path(base_path.ht29,
                                                "crisprn","plasmid_test","counts","combined_counts.txt"), 
                                 rename=T)

plot_sgrna_hist(counts_ps, "HT29; Post-selection")

```


## Kusa v3: sgRNA library

```{r}
kusa.v3 <- suppressMessages(suppressWarnings(read_table2(file.path(base_path.data,"cleanr.tsv")))) %>%
  group_by(GENES) %>%
  summarise(sgRNA_per_gene_count = n()) %>%
  ungroup() %>%
  arrange(GENES) %>%
  rename(gene = GENES) %>%
  filter(gene != "Control" & gene != ".")

plot_sgrna_hist(kusa.v3, "Kusa v3 Original Library")

```

## Summary

```{r}
kable(summarise.sgrna_counts(kusa.v3,"Kusa_v3"))

```

# Orphan sgRNAs

Defined as sgRNAs that don't have gene IDs in the `GENES` column of the original Kusa v3 file and are missing in the Plasmid count files (entered as '.'), but have valid gene IDs in their sgRNA IDs. Several genes in here have all 6 of their sgRNAs present (see sgRNA counts below).

```{r}
orphans <- suppressMessages(suppressWarnings(read_table2(file.path(base_path.data,"cleanr.tsv")))) %>%
  filter(GENES == ".") %>%
  mutate(gene = gsub("^(\\S+?)_.*$","\\1",CODE))

kable(table(orphans$gene))

```

